---
title: "Processing merged seurat"
output: html_notebook
---

# Dependencies
```{r}
library(Seurat)
library(harmony)
library(dplyr)
library(viridis)
```

# Load the data

```{r}
seurat <- readRDS("../output/merged_data_unfiltered.rds")
```

# Pre-processing the data

```{r}
cells_keep <- colnames(seurat)[seurat$Age != "NA"] # removes doublets and negatives
seurat <- subset(seurat, cells=cells_keep)
```

```{r}
seurat[["percent.mt"]] <- PercentageFeatureSet(seurat, pattern = "^mt-")
VlnPlot(seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

## Normalization and feature selection
```{r}
seurat <- NormalizeData(seurat)
seurat <- FindVariableFeatures(seurat, selection.method = "vst",
                               nfeatures=3000)
```

## Cell Cycle Classification
I have run through this without regressing out cell cycle and have found that embeddings and clusterings end up being driven by cell cycle stage, so we'll regress it out.

```{r}
s.genes <- Hmisc::capitalize(tolower(cc.genes$s.genes))
g2m.genes <- Hmisc::capitalize(tolower(cc.genes$g2m.genes))
```

```{r}
seurat <- CellCycleScoring(seurat, s.features = s.genes, g2m.features = g2m.genes, 
                           set.ident = FALSE)
```

## Scaling and regressing out cell cycle

```{r}
seurat <- ScaleData(seurat, features=VariableFeatures(seurat),
                    vars.to.regress=c("percent.mito", "nCount_RNA",
                                      "S.Score", "G2M.Score"))
seurat <- RunPCA(seurat, npcs=100, verbose=F)
ElbowPlot(seurat, ndims=100)
```

## Integrate data with Harmony
Goal here is to remove batch effects across samples

```{r}
seurat$Sample <- paste0(seurat$Age, "_", seurat$Oxygen)
```

```{r}
seurat <- RunHarmony(seurat, "Sample", plot_convergence = T)
```

```{r}
seurat <- RunUMAP(seurat, reduction="harmony", dims=1:40)
```

```{r}
seurat <- FindNeighbors(seurat, reduction='harmony', dims=1:40)
seurat <- FindClusters(seurat, resolution=0.2)
```


```{r}
temp_data <- seurat@meta.data
temp_data$UMAP1 <- Embeddings(seurat, 'umap')[,1]
temp_data$UMAP2 <- Embeddings(seurat, 'umap')[,2]
temp_data$Cluster <- Idents(seurat)
shuffle <- sample(1:ncol(seurat), replace=F)
temp_data <- temp_data[shuffle,]

data_plot <- ggplot(temp_data, aes(x=UMAP1, y=UMAP2)) +
  geom_point(size=0.1, alpha=0.75, aes(color=Oxygen)) +
  theme_void()
ggsave(data_plot, filename = "~/Downloads/umap.oxygen.png", width=9, height=6)
```


```{r}
DimPlot(seurat, label=T)
DimPlot(seurat, group.by="Oxygen")
DimPlot(seurat, group.by="Age")
DimPlot(seurat, group.by="Phase")
```

```{r}
saveRDS(seurat, file="../output/merged_data_filtered.rds")
```

