---
title: "Subset - endothelial"
output: html_notebook
---

# Dependencies

```{r}
library(Seurat)
library(muscat)
library(ggplot2)
library(dplyr)
library(fgsea)
```

# Load the data

```{r}
seurat <- readRDS("../../output/seurat_integrated.rds")
```

# Find endothelial populations

```{r}
DimPlot(seurat, label=T)
FeaturePlot(seurat, features="Cldn5")
```

and subset

```{r}
seurat <- subset(seurat, idents=c(1,10,17))
```

# Re-process
Basic strategy here is to re-align, regress out cell cycle, re-cluster, differential state analysis

For visualization, we'll show non-aligned embeddings, but clusters can be based on aligned data
```{r}
seurat.list <- SplitObject(seurat, split.by="Sample")
```

```{r}
for (i in 1:length(seurat.list)) {
    seurat.list[[i]] <- SCTransform(seurat.list[[i]], 
                                    vars.to.regress=c("S.Score", "G2M.Score",
                                                      "percent.mt"),
                                    verbose = FALSE)
}


seurat.features <- SelectIntegrationFeatures(seurat.list, 
                                             nfeatures = 3000)
seurat.list <- PrepSCTIntegration(seurat.list, 
                                  anchor.features = seurat.features,
                                  verbose = T)
seurat.anchors <- FindIntegrationAnchors(seurat.list, 
                                         normalization.method = "SCT",
                                         anchor.features = seurat.features,
                                         verbose = T)
seurat <- IntegrateData(seurat.anchors, 
                        normalization.method = "SCT",
                        verbose = T)
rm(seurat.anchors, seurat.list, seurat.features)

# Dimensionality reduction and re-cluster
seurat <- RunPCA(seurat, verbose=F)
seurat <- RunUMAP(seurat, dims=1:20)
seurat <- FindNeighbors(seurat, dims=1:20)
seurat <- FindClusters(seurat, resolution=0.25)
```

20PCs is reasonable for embedding -- 30 starts looking a little noisey
```{r}
ElbowPlot(seurat, ndims=50)
```

```{r}
DimPlot(seurat, label=T)
DimPlot(seurat, group.by="Oxygen")
```

Likely some doublets here. Let's look at markers

```{r}
markers <- FindAllMarkers(seurat, only.pos = T, logfc.threshold = 2)
```

```{r}
markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
```

Cluster 5 = fibroblast
Cluster 6 = ??? No unique marker--lots of cell types in the full data express its markers. Definitely not a unique cell type. No markers other than cell cycle. Will remove
Cluster 7 = CD45+
Cluster 8 = AT2 doublets

```{r}
DimPlot(seurat, group.by="Phase")
FeaturePlot(seurat, features=c("Ptprc", "Col1a1", "Sftpc"))
```


# Filter out contaminating populations
```{r}
seurat <- subset(seurat, idents=c(0,1,2,3,4))
```

```{r}
seurat.list <- SplitObject(seurat, split.by="Sample")

for (i in 1:length(seurat.list)) {
    seurat.list[[i]] <- SCTransform(seurat.list[[i]], 
                                    vars.to.regress=c("S.Score", "G2M.Score",
                                                      "percent.mt"),
                                    verbose = FALSE)
}


seurat.features <- SelectIntegrationFeatures(seurat.list, 
                                             nfeatures = 3000)
seurat.list <- PrepSCTIntegration(seurat.list, 
                                  anchor.features = seurat.features,
                                  verbose = T)
seurat.anchors <- FindIntegrationAnchors(seurat.list, 
                                         normalization.method = "SCT",
                                         anchor.features = seurat.features,
                                         verbose = T)
seurat <- IntegrateData(seurat.anchors, 
                        normalization.method = "SCT",
                        verbose = T)
rm(seurat.anchors, seurat.list, seurat.features)

# Dimensionality reduction and re-cluster
seurat <- RunPCA(seurat, verbose=F)
seurat <- RunUMAP(seurat, dims=1:20)
seurat <- FindNeighbors(seurat, dims=1:20)
seurat <- FindClusters(seurat, resolution=0.08)
```

```{r}
DimPlot(seurat, label=T)
DimPlot(seurat, group.by="Oxygen")
```

Clustering looks good. Let's do an un-aligned embedding

# Re-embed unaligned data

```{r}
DefaultAssay(seurat) <- "RNA"
seurat <- SCTransform(seurat, vars.to.regress=c("S.Score", "G2M.Score", "percent.mito"))
seurat <- RunPCA(seurat, verbose=F)
seurat <- RunUMAP(seurat, dims=1:20)
```

```{r}
DimPlot(seurat, label=T)
DimPlot(seurat, group.by="Sample")
DimPlot(seurat, group.by="Oxygen")
```

```{r}
DimPlot(seurat, split.by="Oxygen", group.by="Age", ncol=2)
```


# Save point
```{r}
saveRDS(seurat, "../../output/endothelial_cells.rds")
```


# Cell type frequency
```{r}
clusters_freq <- seurat@meta.data %>%
  group_by(Oxygen, seurat_clusters) %>%
  summarise(n=n()) %>%
  mutate(relative_freq = n/sum(n))
clusters_freq$seurat_clusters <- factor(clusters_freq$seurat_clusters)
clusters_freq$Oxygen <- factor(clusters_freq$Oxygen, levels=c("Normoxia", "Hyperoxia"))
```

```{r}
dark2 <- RColorBrewer::brewer.pal(5, "Dark2")
freq_plot <- ggplot(clusters_freq, aes(x=seurat_clusters, y=relative_freq)) +
  geom_col(aes(fill=Oxygen), color="black", position="dodge") +
  ylab("Relative Frequency") +
  scale_fill_manual(values=dark2[1:2]) +
  scale_y_continuous(expand=c(0,0)) +
  facet_wrap( ~ seurat_clusters, ncol=6, scales="free") +
  theme_classic() +
  theme(strip.background=element_blank(),
        strip.text = element_text(size=11),
        legend.title = element_blank(),
        legend.text = element_text(size=12),
        axis.text.y=element_text(size=10, color='black'),
        axis.text.x=element_text(size=10, color='black', angle=45, hjust=1),
        axis.title.x=element_blank(),
        axis.ticks.x=element_blank())
```

```{r}
freq_plot
```

# Differential state analysis

```{r}
seurat.tmp <- seurat
seurat.tmp$UniqueSample <- paste(seurat.tmp$orig.ident)
# Technically P3_P7_1_Bar1 and P3_P7_2_Bar1, for example, are the same mice--should merge
seurat.tmp$UniqueSample[seurat.tmp$UniqueSample %in% c("P3_P7_1", "P3_P7_2")] <- "P3_P7"
seurat.tmp$UniqueSample <- paste(seurat.tmp$UniqueSample, seurat.tmp$Barcode, sep="_")

DefaultAssay(seurat.tmp) <- "RNA" #can switch back to integrated if desired

seurat.sce <- as.SingleCellExperiment(seurat.tmp)

seurat.sce <- prepSCE(seurat.sce,
                      cluster_id = "seurat_clusters",
                      group_id = "Oxygen",
                      sample_id = "UniqueSample",
                      drop=T)
seurat.sce$group_id <- factor(seurat.sce$group_id, levels=c("Normoxia", "Hyperoxia"))

nk <- length(kids <- levels(seurat.sce$cluster_id))
ns <- length(sids <- levels(seurat.sce$sample_id))
names(kids) <- kids
names(sids) <- sids

pb <- aggregateData(seurat.sce,
                    assay = "counts", fun = "sum",
                    by = c("cluster_id", "sample_id"))

pb$group_id <- factor(pb$group_id, levels=c("Normoxia", "Hyperoxia"))
res <- pbDS(pb, verbose = FALSE)
tmp <- seurat.sce
counts(tmp) <- as.matrix(counts(tmp))
result_table <- resDS(tmp, res, bind = "row", frq = FALSE, cpm = FALSE)
rm(tmp)

#Trim columns we don't really care about
result_table$p_adj.glb__Hyperoxia <- NULL

#Get the frequency of Nox/Hyx cells in a cluster that express a gene
count_mat <- as.matrix(seurat[["RNA"]]@data) > 0

cluster_list <- unique(result_table$cluster_id)

result_table$Normoxia.frq <- 0
normoxia_cells <- colnames(seurat)[seurat$Oxygen == "Normoxia"]
for(i in 1:length(cluster_list)){
  #Which cells to calc freq on
  cluster_cells <- colnames(seurat)[seurat$seurat_clusters == cluster_list[i]]
  test_cells <- intersect(normoxia_cells, cluster_cells)
  #Which rows of result_table to fill
  row_ind <- which(result_table$cluster_id == cluster_list[i])
  frq <- rowSums(count_mat[result_table$gene[row_ind],test_cells]) / length(test_cells)
  result_table$Normoxia.frq[row_ind] <- frq
}

hyperoxia_cells <- colnames(seurat)[seurat$Oxygen == "Hyperoxia"]
result_table$Hyperoxia.frq <- 0
for(i in 1:length(cluster_list)){
  #Which cells to calc freq on
  cluster_cells <- colnames(seurat)[seurat$seurat_clusters == cluster_list[i]]
  test_cells <- intersect(hyperoxia_cells, cluster_cells)
  #Which rows of result_table to fill
  row_ind <- which(result_table$cluster_id == cluster_list[i])
  frq <- rowSums(count_mat[result_table$gene[row_ind],test_cells]) / length(test_cells)
  result_table$Hyperoxia.frq[row_ind] <- frq
}

# Save!
write.csv(result_table, file = "../../output/dsa_results_endothelial_all.csv", row.names = F)
```

# Normoxia: Time-dependent genes
```{r}
seurat.tmp <- seurat
seurat.tmp$UniqueSample <- paste(seurat.tmp$orig.ident)
# Technically P3_P7_1_Bar1 and P3_P7_2_Bar1, for example, are the same mice--should merge
seurat.tmp$UniqueSample[seurat.tmp$UniqueSample %in% c("P3_P7_1", "P3_P7_2")] <- "P3_P7"
seurat.tmp$UniqueSample <- paste(seurat.tmp$UniqueSample, seurat.tmp$Barcode, sep="_")

DefaultAssay(seurat.tmp) <- "RNA" #can switch back to integrated if desired

#Nox cells
cells_keep <- colnames(seurat.tmp)[seurat.tmp$Oxygen=="Normoxia" &
                                 seurat.tmp$Age %in% c("P3", "P14")]
seurat.tmp <- subset(seurat.tmp, cells=cells_keep)
seurat.tmp$Age <- factor(seurat.tmp$Age, levels=c("P3", "P14"))


seurat.sce <- as.SingleCellExperiment(seurat.tmp)

seurat.sce <- prepSCE(seurat.sce,
                      cluster_id = "seurat_clusters",
                      group_id = "Age",
                      sample_id = "UniqueSample",
                      drop=T)

nk <- length(kids <- levels(seurat.sce$cluster_id))
ns <- length(sids <- levels(seurat.sce$sample_id))
names(kids) <- kids
names(sids) <- sids

pb <- aggregateData(seurat.sce,
                    assay = "counts", fun = "sum",
                    by = c("cluster_id", "sample_id"))
pb$group_id <- factor(pb$group_id, levels=c("P3", "P14"))

#Need to manually set model matrix and contrast here for some reason?? Didn't for Nox/Hyx
ei <- metadata(seurat.sce)$experiment_info
mm <- model.matrix(~0 + ei$group_id)
dimnames(mm) <- list(ei$sample_id, levels(ei$group_id))
contrast <- limma::makeContrasts("P14-P3", levels=mm)

res <- pbDS(pb, design=mm, contrast=contrast, verbose = F)
####

tmp <- seurat.sce
counts(tmp) <- as.matrix(counts(tmp))
result_table <- resDS(tmp, res, bind = "row", frq = FALSE, cpm = FALSE)
rm(tmp)

#Trim columns we don't really care about
result_table[,"p_adj.glb__P14-P3"] <- NULL

#Get the frequency of P3/P14 cells in a cluster that express a gene
count_mat <- as.matrix(seurat[["RNA"]]@data) > 0

cluster_list <- unique(result_table$cluster_id)

result_table$P3.frq <- 0
P3_cells <- colnames(seurat)[seurat$Age == "P3" & seurat$Oxygen == "Normoxia"]
for(i in 1:length(cluster_list)){
  #Which cells to calc freq on
  cluster_cells <- colnames(seurat)[seurat$seurat_clusters == cluster_list[i]]
  test_cells <- intersect(P3_cells, cluster_cells)
  #Which rows of result_table to fill
  row_ind <- which(result_table$cluster_id == cluster_list[i])
  frq <- rowSums(count_mat[result_table$gene[row_ind],test_cells]) / length(test_cells)
  result_table$P3.frq[row_ind] <- frq
}

result_table$P14.frq <- 0
P14_cells <- colnames(seurat)[seurat$Age == "P14" & seurat$Oxygen == "Normoxia"]
for(i in 1:length(cluster_list)){
  #Which cells to calc freq on
  cluster_cells <- colnames(seurat)[seurat$seurat_clusters == cluster_list[i]]
  test_cells <- intersect(P14_cells, cluster_cells)
  #Which rows of result_table to fill
  row_ind <- which(result_table$cluster_id == cluster_list[i])
  frq <- rowSums(count_mat[result_table$gene[row_ind],test_cells]) / length(test_cells)
  result_table$P14.frq[row_ind] <- frq
}

write.csv(result_table, file = "../../output/dsa_results_endothelial_nox_development.csv", row.names = F)
```

# Hyperoxia: Time-dependent genes
```{r}
seurat.tmp <- seurat
seurat.tmp$UniqueSample <- paste(seurat.tmp$orig.ident)
# Technically P3_P7_1_Bar1 and P3_P7_2_Bar1, for example, are the same mice--should merge
seurat.tmp$UniqueSample[seurat.tmp$UniqueSample %in% c("P3_P7_1", "P3_P7_2")] <- "P3_P7"
seurat.tmp$UniqueSample <- paste(seurat.tmp$UniqueSample, seurat.tmp$Barcode, sep="_")

DefaultAssay(seurat.tmp) <- "RNA" #can switch back to integrated if desired

#Nox cells
cells_keep <- colnames(seurat.tmp)[seurat.tmp$Oxygen=="Hyperoxia" &
                                 seurat.tmp$Age %in% c("P3", "P14")]
seurat.tmp <- subset(seurat.tmp, cells=cells_keep)
seurat.tmp$Age <- factor(seurat.tmp$Age, levels=c("P3", "P14"))


seurat.sce <- as.SingleCellExperiment(seurat.tmp)

seurat.sce <- prepSCE(seurat.sce,
                      cluster_id = "seurat_clusters",
                      group_id = "Age",
                      sample_id = "UniqueSample",
                      drop=T)

nk <- length(kids <- levels(seurat.sce$cluster_id))
ns <- length(sids <- levels(seurat.sce$sample_id))
names(kids) <- kids
names(sids) <- sids

pb <- aggregateData(seurat.sce,
                    assay = "counts", fun = "sum",
                    by = c("cluster_id", "sample_id"))
pb$group_id <- factor(pb$group_id, levels=c("P3", "P14"))

#Need to manually set model matrix and contrast here for some reason?? Didn't for Nox/Hyx
ei <- metadata(seurat.sce)$experiment_info
mm <- model.matrix(~0 + ei$group_id)
dimnames(mm) <- list(ei$sample_id, levels(ei$group_id))
contrast <- limma::makeContrasts("P14-P3", levels=mm)

res <- pbDS(pb, design=mm, contrast=contrast, verbose = F)
####

tmp <- seurat.sce
counts(tmp) <- as.matrix(counts(tmp))
result_table <- resDS(tmp, res, bind = "row", frq = FALSE, cpm = FALSE)
rm(tmp)

#Trim columns we don't really care about
result_table[,"p_adj.glb__P14-P3"] <- NULL

#Get the frequency of P3/P14 cells in a cluster that express a gene
count_mat <- as.matrix(seurat[["RNA"]]@data) > 0

cluster_list <- unique(result_table$cluster_id)

result_table$P3.frq <- 0
P3_cells <- colnames(seurat)[seurat$Age == "P3" & seurat$Oxygen == "Hyperoxia"]
for(i in 1:length(cluster_list)){
  #Which cells to calc freq on
  cluster_cells <- colnames(seurat)[seurat$seurat_clusters == cluster_list[i]]
  test_cells <- intersect(P3_cells, cluster_cells)
  #Which rows of result_table to fill
  row_ind <- which(result_table$cluster_id == cluster_list[i])
  frq <- rowSums(count_mat[result_table$gene[row_ind],test_cells]) / length(test_cells)
  result_table$P3.frq[row_ind] <- frq
}

result_table$P14.frq <- 0
P14_cells <- colnames(seurat)[seurat$Age == "P14" & seurat$Oxygen == "Hyperoxia"]
for(i in 1:length(cluster_list)){
  #Which cells to calc freq on
  cluster_cells <- colnames(seurat)[seurat$seurat_clusters == cluster_list[i]]
  test_cells <- intersect(P14_cells, cluster_cells)
  #Which rows of result_table to fill
  row_ind <- which(result_table$cluster_id == cluster_list[i])
  frq <- rowSums(count_mat[result_table$gene[row_ind],test_cells]) / length(test_cells)
  result_table$P14.frq[row_ind] <- frq
}

write.csv(result_table, file = "../../output/dsa_results_endothelial_hyx_development.csv", row.names = F)
```

# GSEA on all clusters by logFC

```{r}
result_table <- read.csv("../../output/dsa_results_epithelial_all.csv")
```

```{r}
hallmarks <- fgsea::gmtPathways("~/Data/GeneLists/hallmark.genesets.v6.1.symbols.gmt")
kegg <- fgsea::gmtPathways("~/Data/GeneLists/kegg.genesets.v6.1.symbols.gmt")
go <- fgsea::gmtPathways("~/Data/GeneLists/GOTerms.BP.v6.1.symbols.gmt")
reactome <- fgsea::gmtPathways("~/Data/GeneLists/reactome.genesets.v6.1.symbols.gmt")

gene_sets <- c(hallmarks, kegg, go, reactome)
```

## Loop through clusters

```{r}
runGSEA <- function(cluster){
  results <- filter(result_table, cluster_id == cluster)
  results <- filter(results, Normoxia.frq >= 0.05 |
                      Hyperoxia.frq >= 0.05)
  results <- arrange(results, desc(logFC__Hyperoxia))
  
  cluster_genes <- results$logFC__Hyperoxia
  names(cluster_genes) <- toupper(results$gene)
  
  gsea <- fgsea(pathways = gene_sets,
                stats = cluster_genes,
                minSize=15,
                maxSize=500,
                nproc = 2,
                nperm=100000)
  gsea$cluster <- cluster
  
  return(gsea)
}
```

```{r}
cluster_list <- unique(result_table$cluster_id)

fgsea_results <- lapply(cluster_list, runGSEA)
```

```{r}
fgsea_results <- do.call("rbind", fgsea_results)
fgsea_results <- as.data.frame(fgsea_results)
fgsea_results$leadingEdge <- as.character(fgsea_results$leadingEdge)
```

```{r}
writexl::write_xlsx(fgsea_results, path = "../../output/dsa_gsea_endothelial_results.xlsx")
```
